---
title: "American Community Survey Analysis"
author: "Frank Neugebauer"
date: "December 15, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ggplot2)
# Import the data
comm_sur <- read.csv("acs-14-1yr-s0201.csv", header=TRUE)
```
1. Document the elements in your data including the categories and data types.
    * Id: Factor with 136 levels
    * Id2: Int
    * Geography: Factor with 136 levels
    * PopGroupId: Int
    * POPGROUP.display.label: Factor with 1 level
    * RacesReported: int
    * HSDegree: num
    * BachDegree
2. Report the output from at least the following functions: str(); nrow(); ncol()
```{r}
str(comm_sur)
nrow(comm_sur)
ncol(comm_sur)
```
3. Create a Histogram of the HSDegree variable using the ggplot2 package. Set a bin size for the Histogram.
```{r}
ggplot(comm_sur, aes(x = HSDegree)) + geom_histogram(binwidth=1)
```
4. Include a Title and appropriate X/Y axis labels on your Histogram Plot.
```{r}
ggplot(comm_sur, aes(x = HSDegree)) + geom_histogram(binwidth=1) +   
  ggtitle("Histogram of High School Degree") +
  labs(x = "Count of HS Degrees", y = "Frequency")
```

5. Answer the following questions based on the Histogram produced: 
    a) **Based on what you see in this histogram, is the data distribution 
    unimodal?** The histogram does not appear to be unimodel. While there is
    one dominant hump at 90, there's also other smaller ones at 85 and ~94.
    b) **Is it approximately symmetrical?** No - the histogram is clearly 
    skewed right, with counts between 85 and 95 dominating the graph.
    c) **Is it approximately bell-shaped?** No. Again, there are clear sub-peaks
    at 85 and ~94 and valleys at ~86 and ~93. (There are also smaller peaks
    and valleys from ~76 to ~85.
    d) **Is it approximately normal?** Looking at the historgram as a whole
    (and without any other calculations), I'd say it is not normal; the 
    distribution goes from about 65 - 100 with the majority of degrees
    between 85 and 100, which implies the distribution is skewed towards
    the higher values. 
    e) **If not normal, is the distribution skewed? If so, in which direction? 
    Provide specifics in your responses.** Please refer to 5d for my response
    to this. (Skewed right because the majority of values lies above 80 on a 
    scale from ~60 - ~100).
    f) **Include a normal curve to the Histogram that you plotted.**
```{r}
ggplot(comm_sur, aes(x = HSDegree)) + geom_histogram(binwidth=1) +   
  ggtitle("Histogram of High School Degree") +
  labs(x = "Count of HS Degrees", y = "Frequency") +
  geom_density(aes(y=0.045*..count..), colour="red")
```
    
    I'm not entirely sure this is what is required. The red curve doesn't look
    any more normal than the historgram.
    
5. Continued
    g) **Explain whether a normal distribution can accurately be used as a 
    model for this data.** I do not believe so because again, the data is so
    heavily skewed right.
6. Create a Probability Plot of the HSDegree variable.
```{r}
qqnorm(comm_sur$HSDegree, pch = 1, frame = FALSE)
qqline(comm_sur$HSDegree, col = "steelblue", lwd = 2)
```

The solid line is a reference line of the probability plot.

7. Answer the following questions based on the Probability Plot:
    a) **Based on what you see in this probability plot, is the distribution 
    approximately normal? Explain how you know.** While it looks visually
    'close' to normal, the fact remains that The pattern of data is non-linear. 
    This says to me that the distribution is (still) not normal, providing more
    statistical evidence to the visual conclusion earlier.
    b) **If not normal, is the distribution skewed? If so, in which direction? 
    Explain how you know.** The values are far more closely distributed between
    ~-.5 and 1 and on a scale of -2 to 2, this implies a skew to the right.

8. Now that you have looked at this data visually for normality, you will now 
quantify normality with numbers using the stat.desc() function. Include a 
screen capture of the results produced.
```{r}
library(pastecs)
stat.desc(comm_sur$HSDegree, basic = TRUE, desc = TRUE, norm = TRUE)
```

9. In several sentences provide an explanation of the result produced for skew, kurtosis, and z-scores. In addition, explain how a change in the sample size 
may change your explanation?
    * Skew - the value -1.67 says that the distribution has a long tail in the
    negative direction. It also says that median is larger than the mean, which
    is true (88.70 to 87.63). The significant number of values on the tail pulls
    the mean down. More to the points made previously, the mass of values is
    on the right.
    * Kurtosis - this measure desribes the measure of both tails (unlike skew).
    Stated another way, it's a measure of tailedness. (Sources: https://www.spcforexcel.com/knowledge/basic-statistics/are-skewness-and-kurtosis-useful-statistics and https://en.wikipedia.org/wiki/Kurtosis)
    The value 4.35 shows that HSDegree is leptokurtic (>0) with too many scores
    in the tails (in this case on the left tail).
    * z-scores - generally, according to the Discovering Statistics Using R
    "text", z-score is "the value of an observation expressed in standard
    deviation units." However, the z-score is not computed using stats.desc(). 
    The value can be calculated, however, using z = (X - μ) / σ where z is the 
    z-score, X is the value of the element, μ is the population mean, and σ is 
    the standard deviation. In this case, we have no X, so I'm unsure what is
    expected. 

10. Repeat steps 3 through 9 for the BachDegree variable.

3. Create a Histogram of the BachDegree variable using the ggplot2 package. 
Set a bin size for the Histogram.
```{r}
ggplot(comm_sur, aes(x = BachDegree)) + geom_histogram(binwidth=1)
```

4. Include a Title and appropriate X/Y axis labels on your Histogram Plot.
```{r}
ggplot(comm_sur, aes(x = BachDegree)) + geom_histogram(binwidth=1) +   
  ggtitle("Histogram of Bachelor Degree") +
  labs(x = "Count of Bachelor Degrees", y = "Frequency")
```

5. Answer the following questions based on the Histogram produced: 
    a) **Based on what you see in this histogram, is the data distribution 
    unimodal?** The histogram does not appear to be unimodel. While there is
    one dominant hump at 30, there's also other smaller ones throughout the
    disribution, especially above 30.
    b) **Is it approximately symmetrical?** No - the histogram is clearly 
    skewed left, with counts below 40 dominating the graph.
    c) **Is it approximately bell-shaped?** No. Again, there are clear sub-peaks
    throughout and the peak itself is (fairly far) left of center.
    d) **Is it approximately normal?** Looking at the historgram as a whole
    (and without any other calculations), I'd say it is not normal; the 
    distribution is skewed towards the lower values. 
    e) **If not normal, is the distribution skewed? If so, in which direction? 
    Provide specifics in your responses.** Please refer to 5d for my response
    to this. (Skewed left because the majority of values lie below 40 on a 
    scale from 0 - 60).
    f) **Include a normal curve to the Histogram that you plotted.**
```{r}
ggplot(comm_sur, aes(x = BachDegree)) + geom_histogram(binwidth=1) +   
  ggtitle("Histogram of Bachelor Degree") +
  labs(x = "Count of Bachelor Degrees", y = "Frequency") +
  geom_density(aes(y=0.045*..count..), colour="red")
```
    
    I'm not entirely sure this is what is required. The red curve doesn't look
    any more normal than the historgram, as was the case with HSDegree.
    
5. Continued
    g) **Explain whether a normal distribution can accurately be used as a 
    model for this data.** I do not believe so because again, the data is so
    heavily skewed left and the data is also fairly random looking at all the
    peaks and valleys.
6. Create a Probability Plot of the HSDegree variable.
```{r}
qqnorm(comm_sur$BachDegree, pch = 1, frame = FALSE)
qqline(comm_sur$BachDegree, col = "steelblue", lwd = 2)
```

The solid line is a reference line of the probability plot.

7. Answer the following questions based on the Probability Plot:
    a) **Based on what you see in this probability plot, is the distribution 
    approximately normal? Explain how you know.** The Q-Q plot shows a pretty
    normal distribution, given how close the ideal line the values fall.
    b) **If not normal, is the distribution skewed?** N/A.

8. Now that you have looked at this data visually for normality, you will now 
quantify normality with numbers using the stat.desc() function. Include a 
screen capture of the results produced.
```{r}
library(pastecs)
stat.desc(comm_sur$BachDegree, basic = TRUE, desc = TRUE, norm = TRUE)
```

9. In several sentences provide an explanation of the result produced for skew, kurtosis, and z-scores. In addition, explain how a change in the sample size 
may change your explanation?
    * Skew - the value .328 says that the distribution has a short tail in the
    positive direction. This appears to be a more normal distribution.
    * Kurtosis - the value -0.28 shows that BachDegree is platykurtic (<0) 
    with the values (marginally) flat vis-a-vis the tails.
    * z-scores - generally, according to the Discovering Statistics Using R
    "text", z-score is "the value of an observation expressed in standard
    deviation units." However, the z-score is not computed using stats.desc(). 
    The value can be calculated, however, using z = (X - μ) / σ where z is the 
    z-score, X is the value of the element, μ is the population mean, and σ is 
    the standard deviation. In this case, we have no X, so I'm unsure what is
    expected.
    